---
title: "Kaggle: Predicting Housing Prices"
author: "Lizhizi Cui (lc3268) and Mario Saraiva (mhs2195)"
date: "3/28/2018"
output: 
    html_document:
    number_sections: yes
    toc: yes
    toc_float: yes
---

<center>
`Applied Data Science for Social Sciences Project: Preliminary results`
  
`Lizhizi Cui (lc3268) and Mario Saraiva (mhs2195)`
  
`March 28, 2018`
</center>

<p>
  
  
</p> 

```{r}
knitr::opts_chunk$set(echo = TRUE)
```

We want to accurately predict Sales Price based on other variables in the dataset and we will use our findings from inferential section to create a more robust model. 


#Section I
```{r setup, include=FALSE}
library(glmnet)
library(purrr)
library(tidyr)
library(readr)
library(corrplot)
library(shiny)
library(dplyr)
library(ggplot2)
library(e1071)
library(glmnet)
library(mice)
library(lattice)
library(VIM)
library(caret)
library(stargazer)
library(gam)
library(flam)
library(bartMachine)
library(tree)
library(ISLR)
library(RColorBrewer)
library(ggfortify)
library(scales)
library(stargazer)
library(rsconnect)
library(knitr)
library(gridExtra)
library(scales)
library(Rmisc)
library(ggrepel)
library(randomForest)
library(psych)
library(purrr)
library(tidyr)
library(ggplot2)
library(readr)
library(dplyr)
library(corrplot)
library(RColorBrewer)
library(shiny)
library(ggfortify)
library(pander)

all <- read.csv("/Users/lizhizicui/Desktop/Final-Project---Applied-Data-Science/Data/Raw//train.csv")

set.seed(12345)
in_train <- createDataPartition(y = all$SalePrice, p = 3 / 4, list = FALSE)
str(in_train)
training <- all[ in_train, ] # three fourth of data go to training
testing  <- all[-in_train, ]

all <- rbind(training, testing)
```


##There are 1460 instances of training data and 1460 of test data. Total number of attributes equals 81, of which 36 is quantitative, 43 categorical + Id and SalePrice.##

| Variable      | Quantitative? | Class  |
|---------------|---------------|--------|
| 1stFlrSF      | Yes           | int    |
| 2ndFlrSF      | Yes           | int    |
| 3SsnPorch     | Yes           | int    |
| BedroomAbvGr  | Yes           | int    |
| BsmtFinSF1    | Yes           | int    |
| BsmtFinSF2    | Yes           | int    |
| BsmtFullBath  | Yes           | int    |
| BsmtHalfBath  | Yes           | int    |
| BsmtUnfSF     | Yes           | int    |
| EnclosedPorch | Yes           | int    |
| Fireplaces    | Yes           | int    |
| FullBath      | Yes           | int    |
| GarageArea    | Yes           | int    |
| GarageCars    | Yes           | int    |
| GarageYrBlt   | Yes           | int    |
| GrLivArea     | Yes           | int    |
| HalfBath      | Yes           | int    |
| KitchenAbvGr  | Yes           | int    |
| LotArea       | Yes           | int    |
| LotFrontage   | Yes           | int    |
| LowQualFinSF  | Yes           | int    |
| MasVnrArea    | Yes           | int    |
| MiscVal       | Yes           | int    |
| MoSold        | Yes           | int    |
| MSSubClass    | Yes           | int    |
| OpenPorchSF   | Yes           | int    |
| OverallCond   | Yes           | int    |
| OverallQual   | Yes           | int    |
| PoolArea      | Yes           | int    |
| ScreenPorch   | Yes           | int    |
| TotalBsmtSF   | Yes           | int    |
| TotRmsAbvGrd  | Yes           | int    |
| WoodDeckSF    | Yes           | int    |
| YearBuilt     | Yes           | int    |
| YearRemodAdd  | Yes           | int    |
| YrSold.       | Yes           | int    |
| Alley         | No            | Factor |
| BldgType      | No            | Factor |
| BsmtCond      | No            | Factor |
| BsmtExposure  | No            | Factor |
| BsmtFinType1  | No            | Factor |
| BsmtFinType2  | No            | Factor |
| BsmtQual      | No            | Factor |
| CentralAir    | No            | Factor |
| Condition1    | No            | Factor |
| Condition2    | No            | Factor |
| Electrical    | No            | Factor |
| ExterCond     | No            | Factor |
| Exterior1st   | No            | Factor |
| Exterior2nd   | No            | Factor |
| ExterQual     | No            | Factor |
| Fence         | No            | Factor |
| FireplaceQu   | No            | Factor |
| Foundation    | No            | Factor |
| Functional    | No            | Factor |
| GarageCond    | No            | Factor |
| GarageFinish  | No            | Factor |
| GarageQual    | No            | Factor |
| GarageType    | No            | Factor |
| Heating       | No            | Factor |
| HeatingQC     | No            | Factor |
| HouseStyle    | No            | Factor |
| KitchenQual   | No            | Factor |
| LandContour   | No            | Factor |
| LandSlope     | No            | Factor |
| LotConfig     | No            | Factor |
| LotShape      | No            | Factor |
| MasVnrType    | No            | Factor |
| MiscFeature   | No            | Factor |
| MSZoning      | No            | Factor |
| Neighborhood  | No            | Factor |
| PavedDrive    | No            | Factor |
| PoolQC        | No            | Factor |
| RoofMatl      | No            | Factor |
| RoofStyle     | No            | Factor |
| SaleCondition | No            | Factor |
| SaleType      | No            | Factor |
| Street        | No            | Factor |
| Utilities.    | No            | Factor |

`Quantitative`: 1stFlrSF, 2ndFlrSF, 3SsnPorch, BedroomAbvGr, BsmtFinSF1, BsmtFinSF2, BsmtFullBath, BsmtHalfBath, BsmtUnfSF, EnclosedPorch, Fireplaces, FullBath, GarageArea, GarageCars, GarageYrBlt, GrLivArea, HalfBath, KitchenAbvGr, LotArea, LotFrontage, LowQualFinSF, MSSubClass, MasVnrArea, MiscVal, MoSold, OpenPorchSF, OverallCond, OverallQual, PoolArea, ScreenPorch, TotRmsAbvGrd, TotalBsmtSF, WoodDeckSF, YearBuilt, YearRemodAdd, YrSold.

`Qualitative`: Alley, BldgType, BsmtCond, BsmtExposure, BsmtFinType1, BsmtFinType2, BsmtQual, CentralAir, Condition1, Condition2, Electrical, ExterCond, ExterQual, Exterior1st, Exterior2nd, Fence, FireplaceQu, Foundation, Functional, GarageCond, GarageFinish, GarageQual, GarageType, Heating, HeatingQC, HouseStyle, KitchenQual, LandContour, LandSlope, LotConfig, LotShape, MSZoning, MasVnrType, MiscFeature, Neighborhood, PavedDrive, PoolQC, RoofMatl, RoofStyle, SaleCondition, SaleType, Street, Utilities.

```{r, eda}

hist(all$SalePrice, breaks=seq(30000, 760000, 500),prob=TRUE )
lines(density(all$SalePrice), col="blue", lwd=2)       
lines(density(all$SalePrice, adjust=2), lty="dotted")
summary(all$SalePrice)

# SalePrice does not have a standard normal distribution, it is right skewed

# Log the SalePrice, the distribution of log(SalePrice) looks like a normal distribution
log_saleprice <- log(all$SalePrice)

hist(log_saleprice, breaks=seq(10, 14, 0.025),prob=TRUE )
lines(density(log_saleprice), col="blue", lwd=2)        
```

## Correlation with SalePrice
What numeric variables correlated with SalePrice
```{R, EDA}
all$Id <- rownames(all)
numericVars <- which(sapply(all, is.numeric))
numericVarNames <- names(numericVars)
length(numericVars)


all_numVar <- all[, numericVars]
cor_numVar <- cor(all_numVar, use="pairwise.complete.obs") #correlations of all numeric variables

#sort on decreasing correlations with SalePrice
cor_sorted <- as.matrix(sort(cor_numVar[,'SalePrice'], decreasing = TRUE))
 #select only high corelations
CorHigh <- names(which(apply(cor_sorted, 1, function(x) abs(x)>0.5)))
cor_numVar <- cor_numVar[CorHigh, CorHigh]

corrplot.mixed(cor_numVar, tl.col="black", tl.pos = "lt")

```

According to the correlation table, variable 'Overall Quality'(Overall material and finish quality), the 'GrLivArea' (‘Above Grade’ Living Area) and 'GarageCars'(Size of garage in car capacity) are the top 3 variables that highly correlated to SalePrice.
  
For the rest,it is also clear the multicollinearity is an issue. For example: the correlation between GarageCars and GarageArea is very high (0.89), and both have similar (high) correlations with SalePrice. The other 6 six variables with a correlation higher than 0.5 with SalePrice are: -TotalBsmtSF: Total square feet of basement area -1stFlrSF: First Floor square feet -FullBath: Full bathrooms above grade -TotRmsAbvGrd: Total rooms above grade (does not include bathrooms) -YearBuilt: Original construction date -YearRemodAdd: Remodel date (same as construction date if no remodeling or additions)

## Overall Quality 
  
    
Overall Quality is observed to have positive correlation with SalePrice. It does not seem to have noticeble outliers apart from a point in level 4. 
```{r}
ggplot(data=all[!is.na(all$SalePrice),], aes(x=factor(OverallQual), y=SalePrice))+
        geom_boxplot() + labs(x='Overall Quality') +
        scale_y_continuous(breaks= seq(0, 800000, by=100000), labels = comma)
```

## 'GrLivArea' (‘Above Grade’ Living Area)
  
GrLivArea is observed to have positive correlation with SalePrice, which make sense since big houses are generally more expensive than smaller. One thing worth noticing is that there are two houses with really large living areas and low SalePrice (potential to be outliers). I checked the data and found out they are houses 524, 1299 and 2550(NA for SalePrice).
```{r}
ggplot(data=all[!is.na(all$SalePrice),], aes(x=GrLivArea, y=SalePrice))+
        geom_point(col='black') + geom_smooth(method = "lm", se=FALSE, color="blue", aes(group=1)) +
        scale_y_continuous(breaks= seq(0, 800000, by=100000), labels = comma)
outlier1 <- all%>% filter(GrLivArea > 4500)
outlier1$Id
all[c(524, 1299, 2550), c('SalePrice', 'GrLivArea', 'OverallQual')]

```

## GarageCars
  
The Size of garage shows a positive correlation with SalePrice from level 0-3 but fall at level 4 (non-linear correlation). It make sense because only big houses have large garage, and bigger house usually have higher SalePrice. Also for houses with a garage of capacity of 3 cars(level 3), there are several outliners with much higher SalePrice existing, it must be other factors affecting the outcome price. Something might happen for houses with level 4 garage that needed to explained.
```{r}

summary(all$GarageCars)
ggplot(data=all[!is.na(all$SalePrice),], aes(x=factor(GarageCars), y=SalePrice))+
        geom_boxplot() + labs(x='Size of Garage/car') +
        scale_y_continuous(breaks= seq(0, 800000, by=100000), labels = comma)

outlier2 <- all %>% 
  filter(GarageCars == 3)%>%
  filter(SalePrice > 700000)

outlier2$Id
all[c(524, 1299, 692, 1183), c('SalePrice', 'GrLivArea', 'OverallQual','GarageCars')]
```

##Explore missing values
 
```{r, warning=FALSE, message=FALSE}
na_count <-sapply(all, function(y) sum(length(which(is.na(y)))))
na_count <- data.frame(na_count)
```

Most missing: "PoolQC", "MiscFeature", "Alley", "Fence", "FireplaceQu", "LotFrontage"

Variable | Count of Missing
---------|-----------------
PoolQC | 1453
MiscFeature | 1406
Alley | 1369
Fence | 1179
FireplaceQu | 690
LotFrontage | 259
GarageType | 81
GarageYrBlt | 81
GarageFinish | 81
GarageQual81
GarageCond | 81
BsmtExposur | 38
BsmtFinType2 | 38
BsmtQual | 37
BsmtCond | 37
BsmtFinType1 | 37
MasVnrType | 8

PoolQC: Pool quality
		
       Ex	Excellent
       Gd	Good
       TA	Average/Typical
       Fa	Fair
       NA	No Pool
       
MiscFeature: Miscellaneous feature not covered in other categories
		
       Elev	Elevator
       Gar2	2nd Garage (if not described in garage section)
       Othr	Other
       Shed	Shed (over 100 SF)
       TenC	Tennis Court
       NA	None
       
Alley: Type of alley access to property

       Grvl	Gravel
       Pave	Paved
       NA 	No alley access     
       
Fence: Fence quality
		
       GdPrv	Good Privacy
       MnPrv	Minimum Privacy
       GdWo	Good Wood
       MnWw	Minimum Wood/Wire
       NA	No Fence     
       
FireplaceQu: Fireplace quality

       Ex	Excellent - Exceptional Masonry Fireplace
       Gd	Good - Masonry Fireplace in main level
       TA	Average - Prefabricated Fireplace in main living area or Masonry Fireplace in basement
       Fa	Fair - Prefabricated Fireplace in basement
       Po	Poor - Ben Franklin Stove
       NA	No Fireplace   
       
GarageType: Garage location (**smae for GarageYrBlt/GarageFinish/GarageCars/GarageArea/GarageQual/GarageCond)
		
       2Types	More than one type of garage
       Attchd	Attached to home
       Basment	Basement Garage
       BuiltIn	Built-In (Garage part of house - typically has room above garage)
       CarPort	Car Port
       Detchd	Detached from home
       NA	No Garage    
       
BsmtExposure: Refers to walkout or garden level walls (**smae for BsmtFinType1/BsmtFinSF1/BsmtFinType2/BsmtFinSF2/BsmtUnfSF/BsmtQual)

       Gd	Good Exposure
       Av	Average Exposure (split levels or foyers typically score average or above)	
       Mn	Mimimum Exposure
       No	No Exposure
       NA	No Basement

# Dealing with NA
```{r, PoolQC}
all_test <- all
levels(all_test$PoolQC)

addLevel <- function(x, newlevel=NULL) {
  if(is.factor(x)) {
    if (is.na(match(newlevel, levels(x))))
      return(factor(x, levels=c(levels(x), newlevel)))
  }
  return(x)
}
all_test$PoolQC <- addLevel(all_test$PoolQC, "None")
all_test$PoolQC <- addLevel(all_test$PoolQC, "Po")
all_test$PoolQC <- addLevel(all_test$PoolQC, "TA")
all_test$PoolQC <- ifelse(is.na(all$PoolQC), "None", paste(all$PoolQC))

Qualities <- c('None' = 0, 'Po' = 1, 'Fa' = 2, 'TA' = 3, 'Gd' = 4, 'Ex' = 5)
all_test$PoolQC<- revalue(all_test$PoolQC, Qualities)
all_test$PoolQC <- as.factor(all_test$PoolQC)
```

```{r, MiscFeature}
all_test$MiscFeature <- addLevel(all_test$MiscFeature, "None")
all_test$MiscFeature <- ifelse(is.na(all_test$MiscFeature), "None", paste(all_test$MiscFeature))
all_test$MiscFeature <- as.factor(all_test$MiscFeature)
table(all_test$MiscFeature)
                                                                                            
all_test$Alley <- addLevel(all_test$Alley, "None")
all_test$Alley <- ifelse(is.na(all_test$Alley), "None", paste(all$Alley))
all_test$Alley <- as.factor(all_test$Alley)
table(all_test$Alley)

all_test$Fence <- addLevel(all_test$Fence, "None")
all_test$Fence <- ifelse(is.na(all$Fence), "None", paste(all_test$Fence))
all_test$Fence <- as.factor(all_test$Fence)
table(all_test$Fence)

all_test$FireplaceQu <- addLevel(all_test$FireplaceQu, "None")
all_test$FireplaceQu <- ifelse(!is.na(all_test$FireplaceQu), paste(all_test$FireplaceQu), "None")
all_test$FireplaceQu <- as.factor(all_test$FireplaceQu)


#GarageYrBlt/GarageFinish/GarageCars/GarageArea/GarageQual/GarageCond
all_test$GarageType <- addLevel(all_test$GarageType, "None")
all_test$GarageType <- ifelse(is.na(all_test$GarageType), "None", paste(all_test$GarageType))
all_test$GarageType <- as.factor(all_test$GarageType)
table(all_test$GarageType)


all_test$GarageYrBlt[is.na(all_test$GarageYrBlt)] <- 'No Garage'
all$GarageYrBlt <- as.numeric(all$GarageYrBlt)

all_test$GarageFinish <- addLevel(all_test$GarageFinish, "None")
all_test$GarageFinish <- ifelse(is.na(all_test$GarageFinish), "None", paste(all_test$GarageFinish))
all_test$GarageFinish <- as.factor(all_test$GarageFinish)
table(all_test$GarageFinish)

all_test$GarageCars[is.na(all_test$GarageCars)] <- 'No Garage'
all_test$GarageCars <- as.factor(all_test$GarageCars)
table(all_test$GarageCars)

all_test$GarageArea[is.na(all_test$GarageArea)] <- 'No Garage'
table(all_test$GarageArea)

all_test$GarageQual <- addLevel(all_test$GarageQual, "None")
all_test$GarageQual <- ifelse(is.na(all_test$GarageQual), "None", paste(all_test$GarageQual))
all_test$GarageQual <- as.factor(all_test$GarageQual)
table(all_test$GarageQual)

all_test$GarageCond <- addLevel(all_test$GarageCond, "None")
all_test$GarageCond <- ifelse(is.na(all_test$GarageCond), "None", paste(all_test$GarageCond))
all_test$GarageCond <- as.factor(all_test$GarageCond)
table(all_test$GarageCond)

#BsmtExposure: Refers to walkout or garden level walls (**smae for BsmtFinType1/BsmtFinSF1/BsmtFinType2/BsmtFinSF2/BsmtUnfSF/BsmtQual)
all_test$BsmtQual <- addLevel(all_test$BsmtQual, "None")
all_test$BsmtQual <- ifelse(is.na(all_test$BsmtQual), "None", paste(all_test$BsmtQual))
all_test$BsmtQual <- as.factor(all_test$BsmtQual)
table(all_test$BsmtQual)

all_test$BsmtExposure <- addLevel(all_test$BsmtExposure, "None")
all_test$BsmtExposure <- ifelse(is.na(all_test$BsmtExposure), "None", paste(all_test$BsmtExposure))
all_test$BsmtExposure <- as.factor(all_test$BsmtExposure)
table(all_test$BsmtExposure)

all_test$BsmtFinType1 <- addLevel(all_test$BsmtFinType1, "None")
all_test$BsmtFinType1 <- ifelse(is.na(all_test$BsmtFinType1), "None", paste(all_test$BsmtFinType1))
all_test$BsmtFinType1 <- as.factor(all_test$BsmtFinType1)
str(all_test$BsmtFinType1)

all_test$BsmtCond <- addLevel(all_test$BsmtCond, "None")
all_test$BsmtCond <- ifelse(is.na(all_test$BsmtCond), "None", paste(all_test$BsmtCond))
all_test$BsmtCond <- as.factor(all_test$BsmtCond)
table(all_test$BsmtCond)

all_test$BsmtFinType2 <- addLevel(all_test$BsmtFinType2, "None")
all_test$BsmtFinType2 <- ifelse(is.na(all_test$BsmtFinType2), "None", paste(all_test$BsmtFinType2))
all_test$BsmtFinType2 <- as.factor(all_test$BsmtFinType2)
table(all_test$BsmtFinType2)

all_test$BsmtFinSF1[is.na(all_test$BsmtFinSF1)] <-0
all_test$BsmtFinSF2[is.na(all_test$BsmtFinSF2)] <-0
all_test$BsmtUnfSF[is.na(all_test$BsmtUnfSF)] <-0
all_test$LotFrontage[is.na(all_test$LotFrontage)] <-0
table(all_test$LotFrontage)

drop.na.columns <- c( "MasVnrType","LotFrontage")
all_test<- all_test[ , !(names(all_test) %in% drop.na.columns)]

all_test <- na.omit(all_test)

```
```{r factorize}
all <- all_test
all$Foundation <- as.factor(all$Foundation)
table(all$Foundation)
all$Heating <- as.factor(all$Heating)
all$RoofStyle <- as.factor(all$RoofStyle)
all$RoofMatl <- as.factor(all$RoofMatl)
all$LandContour <- as.factor(all$LandContour)
all$BldgType <- as.factor(all$BldgType)
all$HouseStyle <- as.factor(all$HouseStyle)
all$Condition1 <- as.factor(all$Condition1)
all$Condition2 <- as.factor(all$Condition2)
all$GarageArea <- as.numeric(all$GarageArea)

na_count_test <-sapply(all, function(y) sum(length(which(is.na(y)))))
na_count_test <- data.frame(na_count_test)


write_excel_csv(all, "New_Data.csv")
```

## Finding important variables with Random Forest
The correlation table above provide a good overview of the most important numeric variables and multicolinerity among those variables. Next, I want to get an overview of the most important variables including the categorical variables.

```{r}
# Neighborhood
train <- read.csv("/Users/lizhizicui/Desktop/Final-Project---Applied-Data-Science/Src/New_Data.csv")

train <- na.omit(train)
train$GarageYrBlt <- as.integer(train$GarageYrBlt)
```


```{r}
numericVars <- which(sapply(train, is.numeric)) #index vector numeric variables
factorVars <- which(sapply(train, is.factor)) #index vector factor variables
cat('There are', length(numericVars), 'numeric variables, and', length(factorVars), 'categoric variables')

set.seed(123)
quick_RF <- randomForest(x=train[, -79], y=train$SalePrice, ntree=100,importance=TRUE)
imp_RF <- importance(quick_RF)
imp_DF <- data.frame(Variables = row.names(imp_RF), MSE = imp_RF[,1])
imp_DF <- imp_DF[order(imp_DF$MSE, decreasing = TRUE),]

ggplot(imp_DF[1:20,], aes(x=reorder(Variables, MSE), y=MSE, fill=MSE)) + geom_bar(stat = 'identity') + labs(x = 'Variables', y= '% increase MSE if variable is randomly permuted') + coord_flip() + theme(legend.position="none")
```
####Top 3 most important factors Found !

1) GrLivArea: Above grade (ground) living area square feet
2) OverallQual: Rates the overall material and finish of the house

       10	Very Excellent
       9	Excellent
       8	Very Good
       7	Good
       6	Above Average
       5	Average
       4	Below Average
       3	Fair
       2	Poor
       1	Very Poor
       
3) Neighborhood: Physical locations within Ames city limits

       Blmngtn	Bloomington Heights
       Blueste	Bluestem
       BrDale	Briardale
       BrkSide	Brookside
       ClearCr	Clear Creek
       CollgCr	College Creek
       Crawfor	Crawford
       Edwards	Edwards
       Gilbert	Gilbert
       IDOTRR	Iowa DOT and Rail Road
       MeadowV	Meadow Village
       Mitchel	Mitchell
       Names	North Ames
       NoRidge	Northridge
       NPkVill	Northpark Villa
       NridgHt	Northridge Heights
       NWAmes	Northwest Ames
       OldTown	Old Town
       SWISU	South & West of Iowa State University
       Sawyer	Sawyer
       SawyerW	Sawyer West
       Somerst	Somerset
       StoneBr	Stone Brook
       Timber	Timberland
       Veenker	Veenker

```{r correlation2}
all_numVar <- train[, numericVars]
cor_numVar <- cor(all_numVar, use="pairwise.complete.obs")
cor_sorted <- as.matrix(sort(cor_numVar[,'SalePrice'], decreasing = TRUE))
 #select only high corelations
CorHigh <- names(which(apply(cor_sorted, 1, function(x) abs(x)>0.5)))
cor_numVar <- cor_numVar[CorHigh, CorHigh]

corrplot.mixed(cor_numVar, tl.col="black", tl.pos = "lt", tl.cex = 0.7,cl.cex = .7, number.cex=.7)
```
Since there are several paris of varibles are highly correlated to each other, we can simply drop one of them.
Therefore, I chose to drop 'TotalRmsAbvGrd'(vs GrLivArea); 'GarageArea'(vs GarageCars); 'TotalBsmtSF' (vs X1stFkrSF); 'YearRemodAdd'(vs YearBiult)

```{r drop variables and outliers}
dropVars <- c('YearRemodAdd', 'GarageArea', 'TotalBsmtSF', 'TotalRmsAbvGrd')

train <- train[,!(names(train) %in% dropVars)]

train <- train[-c(524, 1299, 2550),]
write_csv(train, 'Final_data.csv')
```

```{r rf again}
numericVars <- which(sapply(train, is.numeric)) #index vector numeric variables
factorVars <- which(sapply(train, is.factor)) #index vector factor variables
cat('There are', length(numericVars), 'numeric variables, and', length(factorVars), 'categoric variables')

set.seed(123)
quick_RF <- randomForest(x=train[, -76], y=train$SalePrice, ntree=200,importance=TRUE)
imp_RF <- importance(quick_RF)
imp_DF <- data.frame(Variables = row.names(imp_RF), MSE = imp_RF[,1])
imp_DF <- imp_DF[order(imp_DF$MSE, decreasing = TRUE),]

ggplot(imp_DF[1:20,], aes(x=reorder(Variables, MSE), y=MSE, fill=MSE)) + geom_bar(stat = 'identity') + labs(x = 'Variables', y= '% increase MSE if variable is randomly permuted') + coord_flip() + theme(legend.position="none")
```

```{r neighborhood}

df <- data.frame(tapply(train$SalePrice, train$Neighborhood, mean))
library(data.table)
setDT(df, keep.rownames = TRUE)[]
colnames(df) <- c('Neighborhood','Average SalePrice')

ggplot(df, aes(x = Neighborhood, y= `Average SalePrice`)) + 
  geom_col(aes(fill = df$Neighborhood))+
  geom_hline(yintercept=median(train$SalePrice), linetype="dashed", color = "red")+
        theme(axis.text.x = element_text(angle = 45, hjust = 1))

ggplot(data=train, aes(x=Neighborhood)) +
        geom_histogram(stat='count')+
        geom_label(stat = "count", aes(label = ..count.., y = ..count..), size=3)+
        theme(axis.text.x = element_text(angle = 45, hjust = 1))
```


#```{r remove extra levels}
namefactor <- which(sapply(train, is.factor))
DFfactors <- train[, !(names(train) %in% namefactor)]
DFfactors <- DFfactors[, names(DFfactors) != 'SalePrice']
DFdummies <- as.data.frame(model.matrix(~.-1, DFfactors))
dim(DFdummies)

ZerocolTest <- which(colSums(DFdummies[1092:1447,])==0)
colnames(DFdummies[ZerocolTest])
DFdummies <- DFdummies[,-ZerocolTest]

ZerocolTrain <- which(colSums(DFdummies[1:1091,])==0)
colnames(DFdummies[ZerocolTrain])
DFdummies <- DFdummies[,-ZerocolTrain] 
#```

#Modeling
We pick variables that increase MSE if variable is randomly permuted greater than 10% (exclude two variables which are X1ndFlrSF and X2ndFlrSF which representing First/Second floor square feet, I think the information is included in GrLivArea) and run a ols model. The ols model has an (adjusted) R2 that is greater than 0.8
```{r ols}
train1 <- train[1:1092,]
test1 <- train[1092:1447,]

ols <- lm(SalePrice ~ GarageCars+ BsmtFinSF1+ OverallQual+ Neighborhood + GrLivArea ,data = train1)
summary(ols)$adj.r.squared
```
But I don't want to just make my conclusion base on the basis of (possibly flawed) intuition about how real estate prices work.
Thus, I start with a model that includes all available predictors, an interaction between OverallQual and YearBuilt (because the quality of Older houses may not as good as the new ones).
```{r step function}
 # step function 

train1 <- subset(train1, select = -c(Utilities,Street))
test1 <- subset(test1, select = -c(Utilities,Street))
test1 <- test1[!test1$Id == 1004,]
test1 <- test1[!test1$Id == 1299,]
test1 <- test1[!test1$Id == 1001,]
test1 <- test1[!test1$Id == 1188,]
test1 <- test1[!test1$Id == 411,]
test1 <- test1[!test1$Id == 251,]
test1 <- test1[!test1$Id == 596,]

ols <- lm(SalePrice ~ . + OverallQual*YearBuilt - SalePrice,
           data = train1)
ols_AIC <- step(ols, trace = FALSE)
setdiff(names(coef(ols)), names(coef(ols_AIC)))

yhat_AIC <- predict(ols_AIC, newdata = test1)
SSE_AIC <- mean( (test1$SalePrice - yhat_AIC) ^ 2 )
SSE_AIC
```


```{r lasso}
library(lars)
X <- model.matrix(ols)[ ,-1]
y <- train1$SalePrice
lasso <- lars(X, y, type = "lasso", trace = FALSE)

test_X <- model.matrix(ols, data = test1)[ ,-1]
yhat_lasso <- predict(lasso, newx = test_X)$fit
dim(yhat_lasso)

SSE_lasso <- colMeans( (test1$SalePrice - yhat_lasso) ^ 2 )
min(SSE_lasso)

#Trying again with least-angle regression, it finds an equivalent model to be optimal
lar <- lars(X, y, type = "lar", trace = FALSE)
yhat_lar <- predict(lasso, newx = test_X)$fit
dim(yhat_lar)

SSE_lar <- colMeans( (test1$SalePrice - yhat_lar) ^ 2 )
min(SSE_lar) == min(SSE_lasso) 

#Trying again with the forward stagewise algorithm, it also finds an equivalent model to be optimal
fs <- lars(X, y, type = "forward.stagewise", trace = FALSE)
yhat_fs <- predict(lasso, newx = test_X)$fit
dim(yhat_fs)
SSE_fs <- colMeans( (test1$SalePrice - yhat_fs) ^ 2 )
min(SSE_fs) == min(SSE_lasso)


```
 
```{r, warning=FALSE, message=FALSE, echo=FALSE}

mod1_train <- lm(SalePrice ~ (.)^2 + OverallQual*YearBuilt - SalePrice, data = train1)
# stepwise search gets rid of a lot of the interactions
mod1_AIC <- step(mod1_train, trace = FALSE)
names(coef(mod1_AIC))

# Predicting the new model in the testing data set
Yhat_mod1AIC <- predict(mod1_AIC, newdata = testing)

# (Root) MSE
mean( (testing$Outstate - Yhat_mod1AIC) ^ 2 )
sqrt(mean( (testing$Outstate - Yhat_mod1AIC) ^ 2 ))

```

```{r rf again2}
RF <- randomForest(x=train1[, -74], y=train1$SalePrice, ntree=500,importance=TRUE)
RF
plot(RF)
```

Next Steps:
1)Time Series Analysis on Pricing
2)Cluster Analysis on Neighborhood
