---
title: "Kaggle: Predicting Housing Prices"
author: "Lizhizi Cui (lc3268) and Mario Saraiva (mhs2195)"
date: "3/28/2018"
output: 
    html_document:
    number_sections: yes
    toc: yes
    toc_float: yes
runtime: shiny
---

<center>
`Applied Data Science for Social Sciences Project: Preliminary results`
  
`Lizhizi Cui (lc3268) and Mario Saraiva (mhs2195)`
  
`March 28, 2018`
</center>

<p>
  
  
</p> 

```{r}
knitr::opts_chunk$set(echo = TRUE)
```

You could change the working directory. Get the address in the beginning getwd(), replace it by your project folder with setwd(). Then, when accessing a file just use read.table("./folder/file.R")




```{r}
# # install.packages("rstudioapi") # run this if it's your first time using it to install
# library(rstudioapi) # load it
# # the following line is for getting the path of your current open file
# current_path <- getActiveDocumentContext()$path
# # The next line set the working directory to the relevant one:
# setwd(dirname(current_path ))
# # you can make sure you are in the right directory
# print( getwd() )

```

#Section I
```{r setup, include=FALSE}
library(glmnet)
library(purrr)
library(tidyr)
library(readr)
library(corrplot)
library(shiny)
library(dplyr)
library(ggplot2)
library(e1071)
library(glmnet)
library(mice)
library(lattice)
library(VIM)
library(caret)
library(stargazer)
library(gam)
library(flam)
#options(java.parameters = "-Xmx5g") #useful when bartmachine breaks because of Java on IOS
library(bartMachine)
library(tree)
library(ISLR)
library(RColorBrewer)
library(ggfortify)
library(scales)
library(stargazer)
library(rsconnect)
library(knitr)
library(gridExtra)
library(scales)
library(Rmisc)
library(ggrepel)
library(randomForest)
library(psych)
library(purrr)
library(tidyr)
library(ggplot2)
library(readr)
library(dplyr)
library(corrplot)
library(RColorBrewer)
library(shiny)
library(ggfortify)
library(pander)
# setwd(dirname(rstudioapi::getActiveDocumentContext()$path))
#"../data" 

#train <- read.csv("/Volumes/KINGSTON/MDP_2018/Applied Data Science/Final Project/Final-Project---Applied-Data-Science-master/Dataset/Raw/train.csv")

#test <- read.csv("/Volumes/KINGSTON/MDP_2018/Applied Data Science/Final Project/Final-Project---Applied-Data-Science-master/Dataset/Raw/test.csv")

train <- read.csv("/Users/lizhizicui/Desktop/Final-Project---Applied-Data-Science/Data/Raw//train.csv")
test <- read.csv("/Users/lizhizicui/Desktop/Final-Project---Applied-Data-Science/Data/Raw/test.csv")

test$SalePrice <- NA
all_total <- rbind(train,test,stringsAsFactors = FALSE)
all <- all_total
#test <- read.csv('/Users/lizhizicui/Desktop/Dataset/Kaggle House Price/test.csv')

```

##There are 1460 instances of training data and 1460 of test data. Total number of attributes equals 81, of which 36 is quantitative, 43 categorical + Id and SalePrice.##

| Variable      | Quantitative? | Class  |
|---------------|---------------|--------|
| 1stFlrSF      | Yes           | int    |
| 2ndFlrSF      | Yes           | int    |
| 3SsnPorch     | Yes           | int    |
| BedroomAbvGr  | Yes           | int    |
| BsmtFinSF1    | Yes           | int    |
| BsmtFinSF2    | Yes           | int    |
| BsmtFullBath  | Yes           | int    |
| BsmtHalfBath  | Yes           | int    |
| BsmtUnfSF     | Yes           | int    |
| EnclosedPorch | Yes           | int    |
| Fireplaces    | Yes           | int    |
| FullBath      | Yes           | int    |
| GarageArea    | Yes           | int    |
| GarageCars    | Yes           | int    |
| GarageYrBlt   | Yes           | int    |
| GrLivArea     | Yes           | int    |
| HalfBath      | Yes           | int    |
| KitchenAbvGr  | Yes           | int    |
| LotArea       | Yes           | int    |
| LotFrontage   | Yes           | int    |
| LowQualFinSF  | Yes           | int    |
| MasVnrArea    | Yes           | int    |
| MiscVal       | Yes           | int    |
| MoSold        | Yes           | int    |
| MSSubClass    | Yes           | int    |
| OpenPorchSF   | Yes           | int    |
| OverallCond   | Yes           | int    |
| OverallQual   | Yes           | int    |
| PoolArea      | Yes           | int    |
| ScreenPorch   | Yes           | int    |
| TotalBsmtSF   | Yes           | int    |
| TotRmsAbvGrd  | Yes           | int    |
| WoodDeckSF    | Yes           | int    |
| YearBuilt     | Yes           | int    |
| YearRemodAdd  | Yes           | int    |
| YrSold.       | Yes           | int    |
| Alley         | No            | Factor |
| BldgType      | No            | Factor |
| BsmtCond      | No            | Factor |
| BsmtExposure  | No            | Factor |
| BsmtFinType1  | No            | Factor |
| BsmtFinType2  | No            | Factor |
| BsmtQual      | No            | Factor |
| CentralAir    | No            | Factor |
| Condition1    | No            | Factor |
| Condition2    | No            | Factor |
| Electrical    | No            | Factor |
| ExterCond     | No            | Factor |
| Exterior1st   | No            | Factor |
| Exterior2nd   | No            | Factor |
| ExterQual     | No            | Factor |
| Fence         | No            | Factor |
| FireplaceQu   | No            | Factor |
| Foundation    | No            | Factor |
| Functional    | No            | Factor |
| GarageCond    | No            | Factor |
| GarageFinish  | No            | Factor |
| GarageQual    | No            | Factor |
| GarageType    | No            | Factor |
| Heating       | No            | Factor |
| HeatingQC     | No            | Factor |
| HouseStyle    | No            | Factor |
| KitchenQual   | No            | Factor |
| LandContour   | No            | Factor |
| LandSlope     | No            | Factor |
| LotConfig     | No            | Factor |
| LotShape      | No            | Factor |
| MasVnrType    | No            | Factor |
| MiscFeature   | No            | Factor |
| MSZoning      | No            | Factor |
| Neighborhood  | No            | Factor |
| PavedDrive    | No            | Factor |
| PoolQC        | No            | Factor |
| RoofMatl      | No            | Factor |
| RoofStyle     | No            | Factor |
| SaleCondition | No            | Factor |
| SaleType      | No            | Factor |
| Street        | No            | Factor |
| Utilities.    | No            | Factor |

`Quantitative`: 1stFlrSF, 2ndFlrSF, 3SsnPorch, BedroomAbvGr, BsmtFinSF1, BsmtFinSF2, BsmtFullBath, BsmtHalfBath, BsmtUnfSF, EnclosedPorch, Fireplaces, FullBath, GarageArea, GarageCars, GarageYrBlt, GrLivArea, HalfBath, KitchenAbvGr, LotArea, LotFrontage, LowQualFinSF, MSSubClass, MasVnrArea, MiscVal, MoSold, OpenPorchSF, OverallCond, OverallQual, PoolArea, ScreenPorch, TotRmsAbvGrd, TotalBsmtSF, WoodDeckSF, YearBuilt, YearRemodAdd, YrSold.

`Qualitative`: Alley, BldgType, BsmtCond, BsmtExposure, BsmtFinType1, BsmtFinType2, BsmtQual, CentralAir, Condition1, Condition2, Electrical, ExterCond, ExterQual, Exterior1st, Exterior2nd, Fence, FireplaceQu, Foundation, Functional, GarageCond, GarageFinish, GarageQual, GarageType, Heating, HeatingQC, HouseStyle, KitchenQual, LandContour, LandSlope, LotConfig, LotShape, MSZoning, MasVnrType, MiscFeature, Neighborhood, PavedDrive, PoolQC, RoofMatl, RoofStyle, SaleCondition, SaleType, Street, Utilities.

```{R, EDA}

hist(train$SalePrice, breaks=seq(30000, 760000, 500),prob=TRUE )
lines(density(train$SalePrice), col="blue", lwd=2)       
lines(density(train$SalePrice, adjust=2), lty="dotted")
summary(train$SalePrice)

# SalePrice does not have a standard normal distribution, it is right skewed

# Log the SalePrice, the distribution of log(SalePrice) looks like a normal distribution
train$log_saleprice <- log(train$SalePrice)

hist(train$log_saleprice, breaks=seq(10, 14, 0.025),prob=TRUE )
lines(density(train$log_saleprice), col="blue", lwd=2)        
```

## Correlation with SalePrice
What numeric variables correlated with SalePrice
```{R, EDA}
all$Id <- rownames(all)
numericVars <- which(sapply(all, is.numeric))
numericVarNames <- names(numericVars)
length(numericVars)


all_numVar <- all[, numericVars]
cor_numVar <- cor(all_numVar, use="pairwise.complete.obs") #correlations of all numeric variables

#sort on decreasing correlations with SalePrice
cor_sorted <- as.matrix(sort(cor_numVar[,'SalePrice'], decreasing = TRUE))
 #select only high corelations
CorHigh <- names(which(apply(cor_sorted, 1, function(x) abs(x)>0.5)))
cor_numVar <- cor_numVar[CorHigh, CorHigh]

corrplot.mixed(cor_numVar, tl.col="black", tl.pos = "lt")

```

According to the correlation table, variable 'Overall Quality'(Overall material and finish quality), the 'GrLivArea' (‘Above Grade’ Living Area) and 'GarageCars'(Size of garage in car capacity) are the top 3 variables that highly correlated to SalePrice.
  
For the rest,it is also clear the multicollinearity is an issue. For example: the correlation between GarageCars and GarageArea is very high (0.89), and both have similar (high) correlations with SalePrice. The other 6 six variables with a correlation higher than 0.5 with SalePrice are: -TotalBsmtSF: Total square feet of basement area -1stFlrSF: First Floor square feet -FullBath: Full bathrooms above grade -TotRmsAbvGrd: Total rooms above grade (does not include bathrooms) -YearBuilt: Original construction date -YearRemodAdd: Remodel date (same as construction date if no remodeling or additions)

## Overall Quality 
  
    
Overall Quality is observed to have positive correlation with SalePrice. It does not seem to have noticeble outliers apart from a point in level 4. 
```{r}
ggplot(data=all[!is.na(all$SalePrice),], aes(x=factor(OverallQual), y=SalePrice))+
        geom_boxplot() + labs(x='Overall Quality') +
        scale_y_continuous(breaks= seq(0, 800000, by=100000), labels = comma)
```

## 'GrLivArea' (‘Above Grade’ Living Area)
  
GrLivArea is observed to have positive correlation with SalePrice, which make sense since big houses are generally more expensive than smaller. One thing worth noticing is that there are two houses with really large living areas and low SalePrice (potential to be outliers). I checked the data and found out they are houses 524, 1299 and 2550(NA for SalePrice).
```{r}
ggplot(data=all[!is.na(all$SalePrice),], aes(x=GrLivArea, y=SalePrice))+
        geom_point(col='black') + geom_smooth(method = "lm", se=FALSE, color="blue", aes(group=1)) +
        scale_y_continuous(breaks= seq(0, 800000, by=100000), labels = comma)
outlier1 <- all%>% filter(GrLivArea > 4500)
outlier1$Id
all[c(524, 1299, 2550), c('SalePrice', 'GrLivArea', 'OverallQual')]

```

## GarageCars
  
The Size of garage shows a positive correlation with SalePrice from level 0-3 but fall at level 4 (non-linear correlation). It make sense because only big houses have large garage, and bigger house usually have higher SalePrice. Also for houses with a garage of capacity of 3 cars(level 3), there are several outliners with much higher SalePrice existing, it must be other factors affecting the outcome price. Something might happen for houses with level 4 garage that needed to explained.
```{r}

summary(all$GarageCars)
ggplot(data=all[!is.na(all$SalePrice),], aes(x=factor(GarageCars), y=SalePrice))+
        geom_boxplot() + labs(x='Size of Garage/car') +
        scale_y_continuous(breaks= seq(0, 800000, by=100000), labels = comma)

outlier2 <- all %>% 
  filter(GarageCars == 3)%>%
  filter(SalePrice > 700000)

outlier2$Id
all[c(524, 1299, 692, 1183), c('SalePrice', 'GrLivArea', 'OverallQual','GarageCars')]
```

#Section II

##Distribution of Continuous Variables
```{R, Shiny}
#Selecting numeric only and categorical variables.
qt <- train %>% keep(is.numeric)
ql <- train %>% keep(is.factor)

qtc <- na.omit(qt)

qtc$sale_Index <- cut(qtc$SalePrice, 3, labels = c("Minimun.Price.(34.6k.to.275k", "275k.to.515k", "515k.to.7560k"))

table(qtc$sale_Index)

qtc$sale_Index <- as.numeric(qtc$sale_Index)
```

```{r Shiny Continuous Variables, eval=FALSE}

### Shiny uses only numerical variables.
shinyApp(
  ui = fillPage(
    fillCol(flex = c(NA, 1),
            inputPanel(
              selectInput("xcol", label = "X-axis:", 
                          names(qtc))
            ),
            plotOutput("plot", height = "100%")
    )
  ),
  server = function(input, output) {
    output$plot = renderPlot( {
      par(mfrow=c(1,1))
      plot(density(qtc[,input$xcol]), main="Distribution of Quantitative variables", sub=paste("Skewness:", round(e1071::skewness(qtc[,input$xcol]), 2)))
      polygon(density(qtc[,input$xcol]), col="light blue") }
    )},
  options = list(height = 700)
)

```


##Explore missing values
### Version A (Under construction) 
```{r, warning=FALSE, message=FALSE}
na_count <-sapply(all, function(y) sum(length(which(is.na(y)))))
na_count <- data.frame(na_count)
```

Most missing: "PoolQC", "MiscFeature", "Alley", "Fence", "FireplaceQu", "LotFrontage"

Variable | Count of Missing
---------|-----------------
PoolQC | 1453
MiscFeature | 1406
Alley | 1369
Fence | 1179
FireplaceQu | 690
LotFrontage | 259
GarageType | 81
GarageYrBlt | 81
GarageFinish | 81
GarageQual81
GarageCond | 81
BsmtExposur | 38
BsmtFinType2 | 38
BsmtQual | 37
BsmtCond | 37
BsmtFinType1 | 37
MasVnrType | 8

PoolQC: Pool quality
		
       Ex	Excellent
       Gd	Good
       TA	Average/Typical
       Fa	Fair
       NA	No Pool
       
MiscFeature: Miscellaneous feature not covered in other categories
		
       Elev	Elevator
       Gar2	2nd Garage (if not described in garage section)
       Othr	Other
       Shed	Shed (over 100 SF)
       TenC	Tennis Court
       NA	None
       
Alley: Type of alley access to property

       Grvl	Gravel
       Pave	Paved
       NA 	No alley access     
       
Fence: Fence quality
		
       GdPrv	Good Privacy
       MnPrv	Minimum Privacy
       GdWo	Good Wood
       MnWw	Minimum Wood/Wire
       NA	No Fence     
       
FireplaceQu: Fireplace quality

       Ex	Excellent - Exceptional Masonry Fireplace
       Gd	Good - Masonry Fireplace in main level
       TA	Average - Prefabricated Fireplace in main living area or Masonry Fireplace in basement
       Fa	Fair - Prefabricated Fireplace in basement
       Po	Poor - Ben Franklin Stove
       NA	No Fireplace   
       
GarageType: Garage location (**smae for GarageYrBlt/GarageFinish/GarageCars/GarageArea/GarageQual/GarageCond)
		
       2Types	More than one type of garage
       Attchd	Attached to home
       Basment	Basement Garage
       BuiltIn	Built-In (Garage part of house - typically has room above garage)
       CarPort	Car Port
       Detchd	Detached from home
       NA	No Garage    
       
BsmtExposure: Refers to walkout or garden level walls (**smae for BsmtFinType1/BsmtFinSF1/BsmtFinType2/BsmtFinSF2/BsmtUnfSF/BsmtQual)

       Gd	Good Exposure
       Av	Average Exposure (split levels or foyers typically score average or above)	
       Mn	Mimimum Exposure
       No	No Exposure
       NA	No Basement
       

#MasVnrType (Masonry veneer type) and LotFrontage (Linear feet of street connected to property) should not have NA in priciple.  

```{r, PoolQC}
all_test <- all
levels(all_test$PoolQC)

addLevel <- function(x, newlevel=NULL) {
  if(is.factor(x)) {
    if (is.na(match(newlevel, levels(x))))
      return(factor(x, levels=c(levels(x), newlevel)))
  }
  return(x)
}
all_test$PoolQC <- addLevel(all_test$PoolQC, "None")
all_test$PoolQC <- addLevel(all_test$PoolQC, "Po")
all_test$PoolQC <- addLevel(all_test$PoolQC, "TA")
all_test$PoolQC <- ifelse(is.na(all$PoolQC), "None", paste(all$PoolQC))

Qualities <- c('None' = 0, 'Po' = 1, 'Fa' = 2, 'TA' = 3, 'Gd' = 4, 'Ex' = 5)
all_test$PoolQC<- revalue(all_test$PoolQC, Qualities)
table(all_test$PoolQC)
```

```{r, MiscFeature}
all_test$MiscFeature <- addLevel(all_test$MiscFeature, "None")
all_test$MiscFeature <- ifelse(is.na(all$MiscFeature), "None", paste(all$MiscFeature))
all_test$MiscFeature <- as.factor(all_test$MiscFeature)
table(all_test$MiscFeature)

# Neighborhood
train%>%
  group_by(train$Neighborhood)%>%
  mutate(train$avg_price <- mean(train$SalePrice[,train$Neighborhood]))

df <- data.frame(tapply(train$SalePrice, train$Neighborhood, mean))
library(data.table)
setDT(df, keep.rownames = TRUE)[]

ggplot(df, aes(x = df$rn, y= df$tapply.train.SalePrice..train.Neighborhood..mean.)) + geom_col(aes(fill = df$rn))
                                                                                            

all_test$Alley <- addLevel(all_test$Alley, "None")
all_test$Alley <- ifelse(is.na(all$Alley), "None", paste(all$Alley))
all_test$Alley <- as.factor(all_test$Alley)
table(all_test$Alley)

all_test$Fence <- addLevel(all_test$Fence, "None")
all_test$Fence <- ifelse(is.na(all$Fence), "None", paste(all$Fence))
all_test$Fence <- as.factor(all_test$Fence)
table(all_test$Fence)

all_test$FireplaceQu <- addLevel(all_test$FireplaceQu, "None")
all_test$FireplaceQu <- ifelse(!is.na(all$FireplaceQu), paste(all$FireplaceQu), "None")
all_test$FireplaceQu <- as.factor(all_test$FireplaceQu)


#GarageYrBlt/GarageFinish/GarageCars/GarageArea/GarageQual/GarageCond
all_test$GarageType <- addLevel(all_test$GarageType, "None")
all_test$GarageType <- ifelse(is.na(all$GarageType), "None", paste(all$GarageType))
all_test$GarageType <- as.factor(all_test$GarageType)
table(all_test$GarageType)


all_test$GarageYrBlt[is.na(all_test$GarageYrBlt)] <- 'No Garage'
all_test$GarageYrBlt <- as.factor(all_test$GarageYrBlt)
table(all$GarageYrBlt)

all_test$GarageFinish <- addLevel(all_test$GarageFinish, "None")
all_test$GarageFinish <- ifelse(is.na(all$GarageFinish), "None", paste(all$GarageFinish))
all_test$GarageFinish <- as.factor(all_test$GarageFinish)
table(all_test$GarageFinish)

all_test$GarageCars[is.na(all_test$GarageCars)] <- 'No Garage'
all_test$GarageCars <- as.factor(all_test$GarageCars)
table(all_test$GarageCars)

all_test$GarageArea[is.na(all_test$GarageArea)] <- 'No Garage'
all_test$GarageArea <- as.factor(all_test$GarageArea)
table(all_test$GarageArea)

all_test$GarageQual <- addLevel(all_test$GarageQual, "None")
all_test$GarageQual <- ifelse(is.na(all$GarageQual), "None", paste(all$GarageQual))
all_test$GarageQual <- as.factor(all_test$GarageQual)
table(all_test$GarageQual)

all_test$GarageCond <- addLevel(all_test$GarageCond, "None")
all_test$GarageCond <- ifelse(is.na(all$GarageCond), "None", paste(all$GarageCond))
all_test$GarageCond <- as.factor(all_test$GarageCond)
table(all_test$GarageCond)

#BsmtExposure: Refers to walkout or garden level walls (**smae for BsmtFinType1/BsmtFinSF1/BsmtFinType2/BsmtFinSF2/BsmtUnfSF/BsmtQual)
all_test$BsmtQual <- addLevel(all_test$BsmtQual, "None")
all_test$BsmtQual <- ifelse(is.na(all$BsmtQual), "None", paste(all$BsmtQual))
all_test$BsmtQual <- as.factor(all_test$BsmtQual)
table(all_test$BsmtQual)

all_test$BsmtExposure <- addLevel(all_test$BsmtExposure, "None")
all_test$BsmtExposure <- ifelse(is.na(all$BsmtExposure), "None", paste(all$BsmtExposure))
all_test$BsmtExposure <- as.factor(all_test$BsmtExposure)
table(all_test$BsmtExposure)

all_test$BsmtFinType1 <- addLevel(all_test$BsmtFinType1, "None")
all_test$BsmtFinType1 <- ifelse(is.na(all$BsmtFinType1), "None", paste(all$BsmtFinType1))
all_test$BsmtFinType1 <- as.factor(all_test$BsmtFinType1)
table(all_test$BsmtFinType1)

all_test$BsmtFinType2 <- addLevel(all_test$BsmtFinType2, "None")
all_test$BsmtFinType2 <- ifelse(is.na(all$BsmtFinType2), "None", paste(all$BsmtFinType2))
all_test$BsmtFinType2 <- as.factor(all_test$BsmtFinType2)
table(all_test$BsmtFinType2)

all_test$BsmtFinSF1[is.na(all_test$BsmtFinSF1)] <-0
all_test$BsmtFinSF2[is.na(all_test$BsmtFinSF2)] <-0
all_test$BsmtUnfSF[is.na(all_test$BsmtUnfSF)] <-0


drop.na.columns <- c( "MasVnrType","LotFrontage")
all_test<- all_test[ , !(names(all) %in% drop.na.columns)]
```
```{r factorize}
all <- all_test
all$Foundation <- as.factor(all$Foundation)
table(all$Foundation)
all$Heating <- as.factor(all$Heating)
all$RoofStyle <- as.factor(all$RoofStyle)
all$RoofMatl <- as.factor(all$RoofMatl)
all$LandContour <- as.factor(all$LandContour)
all$BldgType <- as.factor(all$BldgType)
all$HouseStyle <- as.factor(all$HouseStyle)
all$Condition1 <- as.factor(all$Condition1)
all$Condition2 <- as.factor(all$Condition2)

all$MoSold <- as.factor(all$MoSold)
all$MSSubClass <- as.factor(all$MSSubClass)

```

## Finding important variables with Random Forest
The correlation table above provide a good overview of the most important numeric variables and multicolinerity among those variables. Next, I want to get an overview of the most important variables including the categorical variables.
```{r}
#set.seed(123)
#quick_RF <- randomForest(x=all.minus.nas[1:1461,-1], y=all.minus.nas$SalePrice[1:1461], ntree=100,importance=TRUE)

```

# Version B (remove all NA by deleting columns)
```{r}
drop.na.columns <- c("PoolQC", "MiscFeature", "Alley", "Fence", "FireplaceQu", "LotFrontage")

ql.minus.nas <- ql[ , !(names(ql) %in% drop.na.columns)]

qlc <- na.omit(ql.minus.nas)
```


##Distribution of Categorical Variables
```{r Shiny Categorical Variables, eval=FALSE}
### Shiny uses only numerical variables.
shinyApp(
  ui = fillPage(
    fillCol(flex = c(NA, 1),
            inputPanel(
              selectInput("xcol", label = "X-axis:", 
                          names(qlc))
            ),
            plotOutput("plot", height = "100%")
    )
  ),
  server = function(input, output) {
    output$plot = renderPlot( {
      par(mfrow=c(1,1))
      plot((qlc[,input$xcol]), main="Distribution of Categorical variables", sub=paste("Skewness:", round(e1071::skewness(qlc[,input$xcol]), 2)), col="light green")
      polygon((qlc[,input$xcol]), col="light blue") }
    )},
  options = list(height = 700)
)
```

#Principal Component Analysis


We do PCA to better understand the main patterns in our data across the whole time period: how the companies differ regarding their financial and ESG score attributes. 
We find an interesting relationship between the Environmental ESG score and financial performance indicators, and see that 2 principal components already explain 40% of the variance.

```{r PCA all QTC Data}
#[,-c(40)]
pca_Cdata <- qtc

pr_out_Cdata <- prcomp(pca_Cdata, scale. = TRUE)

#pr_out_Cdata

#Plot
autoplot(pr_out_Cdata, data = pca_Cdata , colour = "SalePrice", label = TRUE, label.size = 3, shape = FALSE, loadings = TRUE, loadings.colour = 'black',
         loadings.label = TRUE, loadings.label.size = 3, main = "Principal Component Analysis for Continuous Variables)")
par(mar = c(5,4,3,3) + .1, las = 1, cex = 0.8)

pr_var <- pr_out_Cdata$sdev ^ 2
#pr_var
pve <- pr_var / sum(pr_var)
#pve

plot(cumsum(pve), xlab = "Principal Components", ylab = "Cumulative Proportion of Variance Explained", ylim = c(0,1),type = 'b')

```

##PCA and Sales Index

The PCA results are interesting since again, the four ESG variables appear in the opposite direction to the financial variables. It seems that in the short term, ESG and financial performance have a negative relationship.

```{r}
qtc$sale_Index <- cut(qtc$SalePrice, 3, labels = c("Minimun.Price.(34.6k.to.275k", "275k.to.515k", "515k.to.7560k"))

table(qtc$sale_Index)

qtc$sale_Index <- as.numeric(qtc$sale_Index)
```

```{r PCA Sales Index}
df15b <- na.omit(qtc)
# df15 <- df15[-1]

#row.names(df15) <- df15b$Name.1
pca_Cdata15 <- df15b

pr_out_Cdata15 <- prcomp(pca_Cdata15, scale. = TRUE)

#Plot
autoplot(pr_out_Cdata15, data = pca_Cdata15 , colour = "sale_Index", label = TRUE, label.size = 3, shape = FALSE, loadings = TRUE, loadings.colour = 'black',
         loadings.label = TRUE, loadings.label.size = 3, main = "Principal Component Analysis (Sales Index")

par(mar = c(5,4,3,3) + .1, las = 1, cex = 0.8)

pr_var15 <- pr_out_Cdata15$sdev ^ 2
#pr_var
pve15 <- pr_var15 / sum(pr_var15)
#pve

plot(pve15, xlab = "Principal Components", ylab = "Proportion of Variance Explained", ylim = c(0,1),type = 'b')

plot(cumsum(pve15), xlab = "Principal Components", ylab = "Cumulative Proportion of Variance Explained", ylim = c(0,1),type = 'b')

```

```{r random foreast}
set.seed(2018)
quick_RF <- randomForest(x=all[1:1460,-79], y=all$SalePrice[1:1460], ntree=100,importance=TRUE)
imp_RF <- importance(quick_RF)
imp_DF <- data.frame(Variables = row.names(imp_RF), MSE = imp_RF[,1])
imp_DF <- imp_DF[order(imp_DF$MSE, decreasing = TRUE),]

ggplot(imp_DF[1:20,], aes(x=reorder(Variables, MSE), y=MSE, fill=MSE)) + geom_bar(stat = 'identity') + labs(x = 'Variables', y= '% increase MSE if variable is randomly permuted') + coord_flip() + theme(legend.position="none")

```

```{r lasso}
set.seed(1234)
my_control <-trainControl(method="cv", number=5)
lassoGrid <- expand.grid(alpha = 1, lambda = seq(0.001,0.1,by = 0.0005))

lasso_mod <- train(x=train1, y=all$SalePrice[!is.na(all$SalePrice)], method='glmnet', trControl= my_control, tuneGrid=lassoGrid) 
lasso_mod$bestTune

LassoPred <- predict(lasso_mod, test1)
predictions_lasso <- exp(LassoPred) #need to reverse the log to the real values
head(predictions_lasso)

sub_lasso <- data.frame(Id = test_labels, SalePrice = predictions_lasso)

```
Detailed look into PCA results:

`r pander(pr_out_Cdata15)`



