---
title: "Untitled"
author: "LizhiziCui"
date: "2/28/2018"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)

train <- read.csv('/Users/lizhizicui/Desktop/Dataset/Kaggle House Price/train.csv')
test <- read.csv('/Users/lizhizicui/Desktop/Dataset/Kaggle House Price/test.csv')
```


##There are 1460 instances of training data and 1460 of test data. Total number of attributes equals 81, of which 36 is quantitative, 43 categorical + Id and SalePrice.##

Quantitative: 1stFlrSF, 2ndFlrSF, 3SsnPorch, BedroomAbvGr, BsmtFinSF1, BsmtFinSF2, BsmtFullBath, BsmtHalfBath, BsmtUnfSF, EnclosedPorch, Fireplaces, FullBath, GarageArea, GarageCars, GarageYrBlt, GrLivArea, HalfBath, KitchenAbvGr, LotArea, LotFrontage, LowQualFinSF, MSSubClass, MasVnrArea, MiscVal, MoSold, OpenPorchSF, OverallCond, OverallQual, PoolArea, ScreenPorch, TotRmsAbvGrd, TotalBsmtSF, WoodDeckSF, YearBuilt, YearRemodAdd, YrSold

Qualitative: Alley, BldgType, BsmtCond, BsmtExposure, BsmtFinType1, BsmtFinType2, BsmtQual, CentralAir, Condition1, Condition2, Electrical, ExterCond, ExterQual, Exterior1st, Exterior2nd, Fence, FireplaceQu, Foundation, Functional, GarageCond, GarageFinish, GarageQual, GarageType, Heating, HeatingQC, HouseStyle, KitchenQual, LandContour, LandSlope, LotConfig, LotShape, MSZoning, MasVnrType, MiscFeature, Neighborhood, PavedDrive, PoolQC, RoofMatl, RoofStyle, SaleCondition, SaleType, Street, Utilities,

```{R, EDA}

summary(train)
summary(test)
hist(train$SalePrice, breaks=seq(30000, 760000, 500),prob=TRUE )
lines(density(train$SalePrice), col="blue", lwd=2)       
lines(density(train$SalePrice, adjust=2), lty="dotted")


library(ggplot2)


hist(train$SalePrice, breaks=seq(30000, 760000, 2000),prob=TRUE )
lines(density(train$SalePrice), col="blue", lwd=2)       
# indicate not a standard normal distribution

# Log the SalePrice, the distribution of log(SalePrice) looks like a normal distribution
train$log_saleprice <- log(train$SalePrice)
train$log_saleprice
hist(train$log_saleprice, breaks=seq(10, 14, 0.025),prob=TRUE )
lines(density(train$log_saleprice), col="blue", lwd=2)      

library(purrr)
library(tidyr)
library(ggplot2)
library(readr)
library(dplyr)
library(corrplot)
library(RColorBrewer)


is.numeric(train$)
is.factor(train$Electrical)

qt <- train %>% keep(is.numeric)
ql <- train %>% keep(is.factor)

hist(train$log_saleprice,prob=TRUE )
lines(density(train$log_saleprice), col="blue", lwd=2) 

boxplot(SalePrice~BldgType,data=train, main="SalePrince vs Type of dwelling", 
  	xlab="Type of dwelling", ylab="SalePrince")

corrplot(cor(qt), diag = FALSE, method = "number")

col <- brewer.pal(4,"RdYlBu") #selecting color palette
library(RColorBrewer)
corrplot(cor(qt, use = "complete.obs"), col = col, tl.cex = 0.6, number.cex = .6, title = "Correlation between (numeric) Variables", type = "upper", mar = c(1,0,1,0), diag = FALSE, method = "square")

```



```{r cars}
summary(cars)
```



```{r pressure, echo=FALSE}
plot(pressure)
```

